---
title: "visa"
author: "Narges shahmohammadi(yxs662)"
date: '2022-10-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries

```{r cars , echo=FALSE ,include=FALSE }
library(caret)
library(histogram)
library(mlbench)
library(e1071)
library(kernlab)
library(corrplot)
library(pls)
library(elasticnet)
library(mice)
library(VIM)
library(missForest)
library(epiDisplay)
library(pROC) 
library(MASS) 
library(klaR) 
library(mda)  
library(Cubist)
library(pamr)
library(ggplot2)
library(lattice)
library(foreach)
library(splines)
library(gam)
library(knitr)
library(dplyr)
library(moments)
library(creditmodel)
library(ROSE)
library(ROCR)
library(readxl)
library(MLmetrics)
```

# load the data
```{r, echo=FALSE}
setwd("C:/Users/narge/OneDrive/Fall 2022/DATA APPLICATIONS/pro")
visa = read.csv("Visa.csv")
str(visa)
```
* The dataset has 25480 rows and 12 columns 
* `no_of_employees`, ` yr_of_estab`, and `prevailing_wage` are numeric features while rest are objects.
```{r}
visa$case_id <- factor(visa$case_id)
visa$continent <- factor(visa$continent)
visa$education_of_employee <- factor(visa$education_of_employee)
visa$has_job_experience <- factor(visa$has_job_experience)
visa$requires_job_training <- factor(visa$requires_job_training)
visa$region_of_employment <- factor(visa$region_of_employment)
visa$unit_of_wage <- factor(visa$unit_of_wage)
visa$full_time_position <- factor(visa$full_time_position)
visa$case_status <- factor(visa$case_status)
```


```{r}
sum(is.na(visa))
```
* There are no null or missing values in the dataset.

```{r}
sum(duplicated(visa))
```
* There are no duplicate values in the data. 

## EDA
```{r}
#par(mfrow=c(3,4))
hist (visa$no_of_employees,xlab = "no_of_employees", main = "no_of_employees",col='light blue')
hist (visa$yr_of_estab,xlab = "yr_of_estab", main = "yr_of_estab",col='light blue')
hist (visa$prevailing_wage,xlab = "prevailing_wage", main = "prevailing_wage",col='light blue')
```
```{r}
mean(visa$prevailing_wage)
```
#### The median prevailing wage of the employees is around 74k. 

```{r}
mean (visa$no_of_employees)
```
```{r}
unique(visa$no_of_employees)
```

* The distribution of the number of employees is heavily right-skewed.
Some companies have more than 500k employees. Such companies might have multiple offices around the world. The range of the number of employees in a company is huge. There are some anomalies in the data as we can see that the minimum number of employees is equal to -25, which is not possible. I will have to fix this. also its mean is 5667.089 
* The year of establishment of companies ranges from 1800 to 2016, which seems fine.
* The average prevailing wage is 74455.81. There's also a very huge difference in 75th percentile and maximum value which indicates there might be outliers present in this column.


```{r}
sum (visa$no_of_employees< 0)
```
* I will consider the 33 observations as data entry errors and take the absolute values for this column.

```{r}
visa['no_of_employees'] = abs(visa['no_of_employees'])
```

```{r}
unique(visa$no_of_employees)
```

## Let's check the count of each unique category in each of the categorical variables.
```{r}
library(ggplot2)
ggplot(visa,aes(continent))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


### Most of the applications in data are from Asians followed by Europeans.

```{r}
ggplot(visa,aes(education_of_employee))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Most of the applicants have a bachelor's degree followed by a master's degree.

```{r}
ggplot(visa,aes(has_job_experience))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Most of the applicants have job experience and do not require job training.

```{r}
ggplot(visa,aes(	
requires_job_training))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Most of the applicants have job experience and do not require job training.

```{r}
ggplot(visa,aes(	
region_of_employment))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Most applicants have their worksite in the Northeast region of the US.

```{r}
ggplot(visa,aes(	
unit_of_wage))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
* Unit of prevailing wage is an important factor for differentiating between a certified and a denied visa application.
* If the unit of prevailing wage is Yearly, there's a high chance of the application getting certified.


```{r}
ggplot(visa,aes(	
full_time_position))+geom_bar(aes(fill=case_status))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Most of the visa applications are for full-time job positions.

```{r}
ggplot(visa,aes(	
 case_status))+geom_bar(aes(fill=case_status))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### The target column case status is imbalanced with many applicants having a certified visa.
```{r}
visa.numeric <- visa[, unlist(lapply(visa, is.numeric))]
corrplot::corrplot(cor(visa.numeric), method = 'color', order = 'alphabet')
corrplot::corrplot(cor(visa.numeric), method = 'number') 
```
#### there are no correlation between numeric variables
## check the outliers
```{r}
par(mfrow=c(2,2))

boxplot(visa$no_of_employees)
boxplot(visa$prevailing_wage)
boxplot(visa$yr_of_estab)
```
- There are quite a few outliers in the data.
- deleting outliers
```{r}
detect_outlier <- function(x) {
 
    # calculate first quantile
    Quantile1 <- quantile(x, probs=.25)
 
    # calculate third quantile
    Quantile3 <- quantile(x, probs=.75)
 
    # calculate inter quartile range
    IQR = Quantile3-Quantile1
 
    # return true or false
    x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}

# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
 
    # for loop to traverse in columns vector
    for (col in columns) {
 
        # remove observation if it satisfies outlier function
        dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
    }
 
    # return dataframe
    print("Remove outliers")
    print(dataframe)
}


```

```{r}

visa1<-remove_outlier(visa, c('prevailing_wage', 'no_of_employees', 'yr_of_estab')) 
```



```{r}

par(mfrow=c(2,2))

boxplot(visa1$prevailing_wage)
boxplot(visa1$yr_of_estab)
boxplot(visa1$no_of_employees)
```
### it got better...


### Data Preparation for modeling

- I want to predict which visa will be certified.
- I think dont need to case_id column so I will delete it . also, I think that some columns need to be to be dummy 
- I'll split the data into train and test to be able to evaluate the model that I build on the train data.
- finally I will check if the train set is balanced data

```{r}
visa3 <- subset (visa, select = -c(case_id))
visa3
```

```{r}
visa3 <- mutate(visa3,continent_Asia = ifelse(continent=="Asia",1,0))
visa3 <- mutate(visa3,continent_Europe= ifelse(continent=="Europe",1,0))
visa3 <- mutate(visa3,continent_Africa= ifelse(continent=="Africa",1,0))
visa3 <- mutate(visa3,continent_South_America = ifelse(continent=="South America",1,0))
visa3 <- mutate(visa3,continent_North_America = ifelse(continent=="North America",1,0))
visa3 <- mutate(visa3,continent_Oceania = ifelse(continent=="Oceania",1,0))

visa3 <- mutate(visa3,unit_of_wage_Year = ifelse(unit_of_wage=="Year",1,0))
visa3 <- mutate(visa3,unit_of_wage_Week = ifelse(unit_of_wage=="Week",1,0))
visa3 <- mutate(visa3,unit_of_wage_Month = ifelse(unit_of_wage=="Month",1,0))
visa3 <- mutate(visa3,unit_of_wage_Hour = ifelse(unit_of_wage=="Hour",1,0))

visa3 <- mutate(visa3,region_of_employment_Island = ifelse(region_of_employment=="Island",1,0))
visa3 <- mutate(visa3,region_of_employment_West = ifelse(region_of_employment=="West",1,0))
visa3 <- mutate(visa3,region_of_employment_Northeast = ifelse(region_of_employment=="Northeast",1,0))
visa3 <- mutate(visa3,region_of_employment_Midwest = ifelse(region_of_employment=="Midwest",1,0))
visa3 <- mutate(visa3,region_of_employment_South = ifelse(region_of_employment=="South",1,0))

visa3 <- mutate(visa3,education_of_employee_High_School = ifelse(education_of_employee=="High School",1,0))
visa3 <- mutate(visa3,education_of_employee_Doctorate = ifelse(education_of_employee=="Doctorate",1,0))
visa3 <- mutate(visa3,education_of_employee_Bachelor_s = ifelse(education_of_employee=="Bachelor's",1,0))
visa3 <- mutate(visa3,education_of_employee_Master_s = ifelse(education_of_employee=="Master's",1,0))

```

```{r}
set.seed(123)
visa2 <- sample(c(TRUE, FALSE), nrow(visa3), replace=TRUE, prob=c(0.7,0.3))
train <- visa3[visa2, ]
test <- visa3[!visa2, ]
```

FIRST CHECK IF WE HAVE IMBALANCE DATA
```{r}
table(train$case_status)
```

## as you can see we have imbalanced data . I use over sampling to solve this problem after split the data to test and train
## ovun.sample: Over-sampling, under-sampling, combination of over- and under-sampling.
```{r}
train <- ovun.sample(case_status ~ ., data = train, method = "over")$data
table(train$case_status)
```

## now I can run the model#################################


### Model evaluation criterion

### Model can make wrong predictions as:

1. Model predicts that the visa application will get certified but in reality, the visa application should get denied.
2. Model predicts that the visa application will not get certified but in reality, the visa application should get certified. 

### Which case is more important? 
* Both the cases are important as:

* If a visa is certified when it had to be denied a wrong employee will get the job position while US citizens will miss the opportunity to work on that position.

* If a visa is denied when it had to be certified the U.S. will lose a suitable human resource that can contribute to the economy. 

### How to reduce the losses?

* `F1 Score` can be used a the metric for evaluation of the model, greater the F1 score higher are the chances of minimizing False Negatives and False Positives. 
* We will use balanced class weights so that model focuses equally on both classes.

### Decision Tree Model




```{r}
formula = case_status ~ has_job_experience  + requires_job_training +no_of_employees  +yr_of_estab+prevailing_wage+full_time_position  +continent_Asia+continent_Europe+          continent_Africa+continent_South_America+continent_North_America+continent_Oceania+unit_of_wage_Year+unit_of_wage_Week+                unit_of_wage_Month+unit_of_wage_Hour+region_of_employment_Island+region_of_employment_West+region_of_employment_Northeast+region_of_employment_Midwest+region_of_employment_South+education_of_employee_High_School+education_of_employee_Doctorate+education_of_employee_Bachelor_s+education_of_employee_Master_s


```


```{r}

#train<-subset(train,select=-c(unit_of_wage,region_of_employment,education_of_employee,continent))
```

```{r}
library(tree)
library(caret)
tree.model= tree(formula, data=train , method='class')
summary(tree.model)
```

```{r}
plot(tree.model)
text(tree.model,pretty = 0,cex=0.6)
```
I don't think that we need tuning this tree...

```{r}
test <- test[,!names(test) %in% 'case_id']
```

# prediction
```{r}
pred.tree=predict(tree.model,test, type = 'class')
```

```{r}
print(length(test$case_status))
print(length(pred.tree))
```


```{r}
plot(pred.tree, test$case_status)
```

```{r}
caret::confusionMatrix(as.factor(test$case_status),(pred.tree))
```

The model has predicted Denied as Denied (True negative),1492 times and Denied as Certified (False negative), 1778 time.
The model has predicted Certified as Denied (False positive),539 times and Certified as Certified , (True positive)2252 times.
The accuracy of the model is 61% which is not very high.
# F1 SCORE
```{r}
library(MLmetrics)
F1_Score(pred.tree,test$case_status)

```
The F-score, also called the F1-score, is a measure of a model's accuracy on a data set. 
I have F1 score of 0.73 which is not very good. so lets try simple logistic regression.
#  logistic regression

```{r}
Log_reg <- glm(formula , data = train, family = binomial())
summary(Log_reg)
```
First notice that this coefficient is statistically significant (associated with a p-value < 0.05), so these predictors influence on case status:
continentEurope / education_of_employee / has_job_experienceY / region_of_employment /unit_of_wage / full_time_positionY 
```{r}
pscl::pR2(Log_reg)["McFadden"]
```
A value of 0.17 is quite low for McFadden’s R2, which indicates that our model does not fits the data very well and has low predictive power.

# PREDICTION glm
```{r}
test$Pred = predict.glm(Log_reg, newdata = test, type ="response")
```

```{r}
test$case_status  <- factor(test$case_status  )
```

```{r}
test$Prediction = ifelse(test$Pred >=0.5,"Certified","Denied")
test$Prediction  <- factor(test$Prediction )
table(test$Prediction)
```

```{r}
caret::confusionMatrix((test$case_status),as.factor(test$Prediction))

```
Accuracy of glm is 28% and it is very low

```{r}
F1_Score(test$Prediction,test$case_status)
```
F1 score is also very low. so we can say this model is not good at all


## randomForest model
```{r }
library(randomForest)
random_Forest <- randomForest(formula, data= train) 
```

```{r}
p1 <- predict(random_Forest, test)
confusionMatrix(p1, test$case_status)
```
Accuracy of this model is 0.72. this result is better than others
```{r}
F1_Score(p1,test$case_status)

```

also its F1 score is good

```{r}
random_Forest$importance
```
## tuning RandomForest
```{r , echo=FALSE ,include=FALSE }
library(randomForestSRC)
library(mlbench)
library(caret)
library(e1071)

set.seed(123)

train$case_status <- as.factor(train$case_status)

rf.forest1 <- rfsrc(formula,data = train, do.trace=T,
                 importance = TRUE,
                 ntree = 1000)

```

## prediction

```{r}
predicted_tunedrf <- predict(rf.forest1,test)

print(predicted_tunedrf)
```

* Misclassification Rate: It tells you what fraction of predictions were incorrect; its formula is : (1-Accuracy) so we can say that the accuracy is about %72 which is pretty close to simple random forest model, but with comparison between their confusion metrics it seems that tuned random model is better 

* Tuned Random Forest model has given a good and generalized performance. We will use it as our final model.

## Compute variable importance to detect interactions and optimize hyper- parameters for acquired customers.

```{r}
rf.forest1$importance
```


* Looking at the feature importance of the Random Forest model, the top 4 important features to look for while certifying a visa are Education of the employee, Job experience , no_of_employees and prevailing_wage. 

### Recommendations

###################

**The profile of the applicants for whom the visa status can be approved:**

**Primary information to look at:**
- Education level - At least has a Bachelor's degree - Master's and doctorate are preferred.
- Job Experience - Should have some job experience. 
- Prevailing wage - The median prevailing wage for the certified applications is slightly higher as compared to denied applications. 


**Secondary information to look at:**

- Unit of Wage- Applicants having a yearly unit of wage. 
- Continent - Ideally the nationality and ethnicity of an applicant shouldn't matter to work in a country but previously it has been observed that applicants from Europe, and Asia have higher chances of visa certification.
- Region of employment - the analysis suggests that the applications to work in the Mid-West region have more chances of visa approval. The approvals can also be made based on requirement of talent, from our analysis we see that:
    * The requirement for the applicants who have passed high school is most in the South region, followed by Northeast region.
    * The requirement for Bachelor's is mostly in South region, followed by West region.
    * The requirement for Master's is most in Northeast region, followed by South region.
    * The requirement for Doctorate's is mostly in West region, followed by Northeast region.
    
    
**The profile of the applicants for whom the visa status can be denied:**

**Primary information to look at:**
- Education level - Doesn't have any degree 
- Job Experience - Doesn't have any job experience. 
- Prevailing wage - The median prevailing wage of the employees for whom the visa got certified is around 65k. 


**Secondary information to look at:**

- 'Unit of Wage' - Applicants having an hourly unit of wage. 
- 'Continent' - Ideally the nationality and ethnicity of an applicant shouldn't matter to work in a country but previously it has been observed that applicants from South America, North America, and Oceania have higher chances of visa applications getting denied.